---
title: "hebstr-doc"
subtitle: "Multi-format Quarto extension"
author: "hebstr"
date: last-modified
format:
  hebstr-doc-html:
    code-tools:
      source: https://github.com/hebstr/quarto-hebstr-doc
---

# Typography and headings

Only `hebstr-doc-html` ships on `main`; Typst and DOCX land once validated.

Body text is [**Luciole**](https://luciole-vision.com) (sans-serif, low-vision).
Inline `code()` is [**Fira Code**](https://github.com/tonsky/FiraCode) with ligatures.

## Level 2 heading

Headings are numbered down to level 4 (`number-depth: 4`).
The cross-reference title delimiter is a literal dot (e.g. *Figure 1.1*).

### Level 3 heading

Paragraph with *emphasis*, **strong emphasis**, a [link to the Quarto docs](https://quarto.org), and an internal reference to @sec-figure.

#### Level 4 heading

Deepest numbered level.

# Lists

Unordered:

- First item
- Second item with a *nested* block:
  - Nested bullet one
  - Nested bullet two
- Third item

Ordered:

1. Initialise the R session
2. Load the tidyverse
3. Render the document

# Figure {#sec-figure}

@fig-penguins is **ggplot2** on `palmerpenguins`, under the shared knitr options (`echo`, `message`, `warning` all off).

The chunk emits two plots; `renderings: [light, dark]` tags them `.light-content` / `.dark-content`, and the body class picks one at runtime.
`renderings` rejects `fig-cap` and `fig-`-prefixed labels as cell options, so the fenced div carries the crossref id and caption.
A transparent device background (`dev.args`) shows the page through, so only the ink varies per mode.

::: {#fig-penguins}
```{r}
#| label: penguins-fig
#| renderings: [light, dark]
#| fig-height: 5
#| dev: svglite
#| dev.args: !expr list(bg = "transparent")

library(ggplot2)

theme_hebstr <- function(ink, grid) {
  theme_minimal(base_size = 12, ink = ink, paper = NA) +
    theme(
      axis.text = element_text(colour = ink),
      panel.grid = element_line(colour = grid),
      axis.ticks = element_line(colour = grid)
    )
}

p <- ggplot(
  palmerpenguins::penguins,
  aes(x = flipper_length_mm, y = bill_length_mm, colour = species)
) +
  geom_point(size = 3, alpha = 0.85) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
  labs(
    x = "Flipper length (mm)",
    y = "Bill length (mm)",
    colour = "Species"
  )

p + theme_hebstr(ink = "#1a1a1a", grid = "#d9d9d9")
p + theme_hebstr(ink = "#e8e8e8", grid = "#4d4d4d")
```

Bill length vs flipper length by species (palmerpenguins).
:::

Both renderings stay in the DOM: mode switches need no reload, and the lightbox tracks the visible one.

# Table

A simple aggregation rendered as a Markdown table.
`tbl-cap-location: top` puts the caption above the table; `tbl-title: "Table"` stays language-neutral.

```{r}
#| label: penguins-count
#| echo: true
#| code-fold: false

library(dplyr)

palmerpenguins::penguins |>
  count(species) |>
  rename(Sp = species)
```

# Code block

`echo: true` overrides the default to exercise highlighting, folding, and the copy-on-hover button (HTML), plus the `code.tmTheme` TextMate theme Typst applies via `#set raw(theme: …)` in `template.typ`.

```{r}
#| label: penguins-summary
#| echo: true
#| eval: false

library(dplyr)

palmerpenguins::penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species, island) |>
  summarise(
    n = n(),
    mean_mass_g = mean(body_mass_g),
    .groups = "drop"
  ) |>
  arrange(desc(mean_mass_g))
```

# Code window

These blocks exercise the embedded [`mcanouil/quarto-code-window`](https://github.com/mcanouil/quarto-code-window) filter.
The extension sets `style: default` (plain filename header, no window chrome); with `auto-filename: true`, blocks without a `filename=` show the language name as the label.

```python
def greet(name: str) -> str:
    return f"Hello, {name}!"
```

```r
penguins |>
  dplyr::count(species)
```

```bash
quarto add hebstr/quarto-hebstr-doc --no-prompt
quarto render example.qmd
```

# Injected script

The `script` shortcode (in-tree `filters/add-code-files.lua`) injects `scripts/demo_penguins.R` verbatim.
It emits a code block with the `cell-code` class, so `code-fold` wraps it in a collapsed `<details>`; the filename header expands it.
`filters/add-code-files.js` rewrites the `<summary>` text to the filename (from [`shafayetShafee/add-code-files`](https://github.com/shafayetShafee/add-code-files), MIT).

Inline chunks and externalised scripts both render as `code-window`, scripts folded behind their filename.

{{< script scripts/demo_penguins.R >}}

# Tabset {#sec-tabs}

Tabs use the `.panel-tabset` div class; `tab-surface` (`theme-base.scss`) sets the panel fill, a tint of `primary` that covers both modes with one value.

::: {.panel-tabset}
## Adelie

```{r}
#| label: penguins-adelie
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::filter(species == "Adelie") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```

## Chinstrap

```{r}
#| label: penguins-chinstrap
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::filter(species == "Chinstrap") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```

## Gentoo

```{r}
#| label: penguins-gentoo
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::filter(species == "Gentoo") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```
:::

# File tree

The `filetree` shortcode (in-tree `filters/filetree.lua`) walks this repository from disk at render time.
`mode` picks the rendering: `static` (default) collapses directories past `depth` to `…`; `dynamic` (here) expands and collapses folders on click, with `depth` setting the open depth on load.

Descriptions come from the `filetree.yml` sidecar, keyed by path.
An unused key raises a render warning naming it, whether the path is absent from disk or dropped by `exclude`, `hidden` or `depth`.

{{< filetree mode=dynamic >}}

HTML renders a nested list styled by the `.filetree` rules in `theme-base.scss`.
In `dynamic` mode each folder is a native `<details>` that toggles without JavaScript, its Material icon switching to the open-folder variant while expanded.
The surface stays dark in both colour modes, like the code windows.
Each entry resolves a Material Icon Theme icon by exact filename, then extension, then a generic fallback: `README.md` matches on name, `CHANGELOG.md` falls through to the Markdown icon, `example.qmd` matches on extension.
Typst and DOCX receive a plain bullet list.

# Equation

Inline math: $y = \beta_0 + \beta_1 x + \varepsilon$.

Display math with a cross-reference target (@eq-ols):

$$
  \hat{\beta} = (X^\top X)^{-1} X^\top y
$$ {#eq-ols}

Textual equations (@eq-text):

$$
  \frac{\text{upper text}}{\text{lower text}} \text{ beside text}
$$ {#eq-text}

# Footnote

Footnotes collect at the document end under a `.footnotes` section.[^footnote-demo]

[^footnote-demo]: The footnote body.
    Markdown works here: *emphasis*, `code()`, and [links](https://quarto.org).

# Callouts

Callouts use `callout-appearance: simple`.

::: {.callout-note}
# A note callout.

{{< lipsum 1 >}}
:::

::: {.callout-tip}
# A tip callout.

{{< lipsum 1 >}}
:::

::: {.callout-caution}
# A caution callout.

{{< lipsum 1 >}}
:::

::: {.callout-warning}
# A warning callout.

{{< lipsum 1 >}}
:::

::: {.callout-important}
# An important callout.

{{< lipsum 1 >}}
:::

# Block quote

> A citation.
>
> --- author

# Cross-references

- Section reference: @sec-figure
- Figure reference: @fig-penguins
- Equation reference: @eq-ols

\ 

# Session {.unnumbered .unlisted}

```{r}
#| label: session-info
#| echo: true

sessioninfo::session_info(pkgs = "attached")
```
