---
title: "hebstr-doc"
subtitle: "Multi-format Quarto extension"
author: "hebstr"
date: last-modified
format:
  hebstr-doc-html:
    code-tools:
      source: https://github.com/hebstr/quarto-hebstr-doc
---

```{r}
#| label: setup
#| include: false

source("_extensions/hebstr-doc/fonts/register.R")
```

# Typography and headings

Only `hebstr-doc-html` ships on `main`; Typst and DOCX land once validated.

Body text is [**Luciole**](https://luciole-vision.com) (sans-serif, low-vision).
Inline `code()` is [**Fira Code**](https://github.com/tonsky/FiraCode) with ligatures.

## Level 2 heading

Headings are numbered down to level 4 (`number-depth: 4`).
The cross-reference title delimiter is a literal dot (e.g. *Figure 1.1*).

### Level 3 heading

Paragraph with *emphasis*, **strong emphasis**, a [link to the Quarto docs](https://quarto.org), and an internal reference to @sec-figure.

#### Level 4 heading

Deepest numbered level.

# Lists

Unordered:

- First item
- Second item with a *nested* block:
  - Nested bullet one
  - Nested bullet two
- Third item

Ordered:

1. Initialise the R session
2. Load the tidyverse
3. Render the document

# Figure {#sec-figure}

@fig-penguins is **ggplot2** on `palmerpenguins`, under the shared knitr options (`echo`, `message`, `warning` all off) and the HTML format's own `dev: svglite`, which Typst and DOCX do not carry.

The chunk emits two plots; `renderings: [light, dark]` tags them `.light-content` / `.dark-content`, and the body class picks one at runtime.
`renderings` rejects `fig-cap` and `fig-`-prefixed labels as cell options, so the fenced div carries the crossref id and caption.
A transparent device background (`dev.args`) shows the page through, so only the ink varies per mode.
Setting `dev.args` replaces the format's own value rather than extending it, which drops the `system_fonts` alias that would otherwise supply the document font, so the theme names `base_family` itself.

::: {#fig-penguins}
```{r}
#| label: penguins-fig
#| renderings: [light, dark]
#| fig-height: 5
#| dev.args: !expr list(bg = "transparent")

library(ggplot2)

theme_hebstr <- function(ink, grid) {
  theme_minimal(
    base_size = 12,
    base_family = "Luciole",
    ink = ink,
    paper = NA
  ) +
    theme(
      axis.text = element_text(colour = ink),
      panel.grid = element_line(colour = grid),
      axis.ticks = element_line(colour = grid)
    )
}

p <- ggplot(
  palmerpenguins::penguins,
  aes(x = flipper_length_mm, y = bill_length_mm, colour = species)
) +
  geom_point(size = 3, alpha = 0.85) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
  labs(
    x = "Flipper length (mm)",
    y = "Bill length (mm)",
    colour = "Species"
  )

p + theme_hebstr(ink = "#1a1a1a", grid = "#d9d9d9")
p + theme_hebstr(ink = "#e8e8e8", grid = "#4d4d4d")
```

Bill length vs flipper length by species (palmerpenguins).
:::

Both renderings stay in the DOM: mode switches need no reload, and the lightbox tracks the visible one.

# Table

`tbl-cap-location: top` puts a caption above its table; `tbl-title: "Table"` stays language-neutral.
Only the `gt` table below carries one.

## Markdown table

A simple aggregation rendered as a Markdown table.

```{r}
#| label: penguins-count
#| echo: true
#| code-fold: false

library(dplyr)

palmerpenguins::penguins |>
  count(species) |>
  rename(Sp = species)
```

## gt

`gt` resolves a table's palette in R and writes it into a `<style>` block scoped by the table's generated id, so a table renders on the colours its code resolved whatever scheme the page is in.
@tbl-penguins-gt asks for a deliberate light palette: a pale blue table ground with white striped rows, bold column labels over a rule, and no vertical hairlines.
The blue is `#F0FAFF`, a step off the `#F2FAFF` that `primary-back` compiles to in light.

Dark is where the one scheme-specific `scss:rules` region of `theme-dark.scss` takes over and replaces it wholesale.
No colour written into that `<style>` block survives, and the two surfaces it replaces them with are the page's own pair: what the light table leaves on the page background (column labels, striped rows, the source note) lands on `primary-surface`, and the table ground on `primary-back`.

```{r}
#| label: tbl-penguins-gt
#| tbl-cap: "Bill length and body mass by species and island."
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species, island) |>
  summarise(
    n = n(),
    bill = round(mean(bill_length_mm, na.rm = TRUE), 1),
    mass = round(mean(body_mass_g)),
    .groups = "drop"
  ) |>
  gt::gt() |>
  gt::cols_label(
    species = "Species",
    island = "Island",
    n = "n",
    bill = "Bill length (mm)",
    mass = "Body mass (g)"
  ) |>
  gt::tab_spanner(label = "Mean", columns = c(bill, mass)) |>
  gt::cols_align(align = "left", columns = c(species, island)) |>
  gt::cols_align(align = "center", columns = c(n, bill, mass)) |>
  gt::tab_source_note("palmerpenguins, records with a body mass.") |>
  gt::tab_options(
    table.width = gt::pct(80),
    table.background.color = "#F0FAFF",
    row.striping.include_table_body = TRUE,
    row.striping.background_color = "white",
    column_labels.background.color = "white",
    column_labels.font.weight = "bold",
    column_labels.border.top.style = "none",
    column_labels.border.bottom.color = "#333333",
    source_notes.background.color = "white",
    table.font.color = "#333333",
    table.border.top.style = "none",
    table.border.bottom.style = "none",
    table_body.border.top.color = "#333333",
    table_body.border.bottom.color = "#333333",
    table_body.hlines.style = "none",
    data_row.padding = gt::px(8)
  )
```

# Code block

`echo: true` overrides the default to exercise highlighting, folding, and the copy-on-hover button (HTML), plus the `code.tmTheme` TextMate theme Typst applies via `#set raw(theme: …)` in `template.typ`.

```{r}
#| label: penguins-summary
#| echo: true
#| eval: false

library(dplyr)

palmerpenguins::penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species, island) |>
  summarise(
    n = n(),
    mean_mass_g = mean(body_mass_g),
    .groups = "drop"
  ) |>
  arrange(desc(mean_mass_g))
```

# Code window

These blocks exercise the embedded [`mcanouil/quarto-code-window`](https://github.com/mcanouil/quarto-code-window) filter.
The extension sets `style: default` (plain filename header, no window chrome); with `auto-filename: true`, blocks without a `filename=` show the language name as the label.

```python
def greet(name: str) -> str:
    return f"Hello, {name}!"
```

```r
penguins |>
  dplyr::count(species)
```

```bash
quarto add hebstr/quarto-hebstr-doc --no-prompt
quarto render example.qmd
```

# Injected script

The `script` shortcode (in-tree `filters/add-code-files.lua`) injects `scripts/demo_penguins.R` verbatim.
It emits a code block with the `cell-code` class, so `code-fold` wraps it in a collapsed `<details>`; the filename header expands it.
`filters/add-code-files.js` rewrites the `<summary>` text to the filename (from [`shafayetShafee/add-code-files`](https://github.com/shafayetShafee/add-code-files), MIT).

Inline chunks and externalised scripts both render as `code-window`, scripts folded behind their filename.

{{< script scripts/demo_penguins.R >}}

# Tabset {#sec-tabs}

Tabs use the `.panel-tabset` div class; `tab-surface` (`theme-base.scss`) sets the panel fill, a tint of `primary` that covers both modes with one value.

::: {.panel-tabset}
## Adelie

```{r}
#| label: penguins-adelie
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::filter(species == "Adelie") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```

## Chinstrap

```{r}
#| label: penguins-chinstrap
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::filter(species == "Chinstrap") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```

## Gentoo

```{r}
#| label: penguins-gentoo
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::filter(species == "Gentoo") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```
:::

# File tree

The `filetree` shortcode (in-tree `filters/filetree.lua`) walks this repository from disk at render time.
`mode` picks the rendering: `static` (default) collapses directories past `depth` to `…`; `dynamic` (here) expands and collapses folders on click, with `depth` setting the open depth on load.

Descriptions come from the `filetree.yml` sidecar, keyed by path.
An unused key raises a render warning naming it, whether the path is absent from disk or dropped by `exclude`, `hidden` or `depth`.

{{< filetree mode=dynamic >}}

HTML renders a nested list styled by the `.filetree` rules in `theme-base.scss`.
In `dynamic` mode each expandable folder is a native `<details>` that toggles without JavaScript, its Material icon switching to the open-folder variant while expanded; a childless folder renders flat.
The surface stays dark in both colour modes, like the code windows.
Each entry resolves a Material Icon Theme icon by exact filename, then extension, then a generic fallback: `README.md` matches on name, `CHANGELOG.md` falls through to the Markdown icon, `example.qmd` matches on extension.
Typst and DOCX receive a plain bullet list.

# Equation

Inline math: $y = \beta_0 + \beta_1 x + \varepsilon$.

Display math with a cross-reference target (@eq-ols):

$$
  \hat{\beta} = (X^\top X)^{-1} X^\top y
$$ {#eq-ols}

Textual equations (@eq-text):

$$
  \frac{\text{upper text}}{\text{lower text}} \text{ beside text}
$$ {#eq-text}

# Footnote

Footnotes collect at the document end under a `.footnotes` section.[^footnote-demo]

[^footnote-demo]: The footnote body.
    Markdown works here: *emphasis*, `code()`, and [links](https://quarto.org).

# Callouts

Callouts use `callout-appearance: simple`.

::: {.callout-note}
# A note callout.

{{< lipsum 1 >}}
:::

::: {.callout-tip}
# A tip callout.

{{< lipsum 1 >}}
:::

::: {.callout-caution}
# A caution callout.

{{< lipsum 1 >}}
:::

::: {.callout-warning}
# A warning callout.

{{< lipsum 1 >}}
:::

::: {.callout-important}
# An important callout.

{{< lipsum 1 >}}
:::

# Block quote

> A citation.
>
> --- author

# Annexes

Quarto ships no appendix prefix among its crossref types, so the extension declares a custom `anx` float (`crossref: custom:` in `_extension.yml`) and reaches it from a chunk through `filters/crossref-anx.lua`.
The knitr engine builds a float only from a `fig-` or `tbl-` label, so an annexe is authored as `tbl-anx-…` or `fig-anx-…`, and the filter drops that carrier prefix at `pre-quarto`.
What is left is a float referenced as `@anx-…` and captioned *Annexe n.*; the three forms below share one counter.

```{r}
#| label: tbl-anx-mass
#| tbl-cap: "Mean body mass by species and island."
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::group_by(species, island) |>
  dplyr::summarise(
    n = dplyr::n(),
    mean_mass_g = round(mean(body_mass_g, na.rm = TRUE)),
    .groups = "drop"
  )
```

The carrier also decides which caption key is read: `tbl-cap` above, `fig-cap` below.

```{r}
#| label: fig-anx-mass-spread
#| fig-cap: "Body mass spread by species."
#| fig-height: 3.5
#| echo: true
#| code-fold: false

ggplot2::ggplot(
  palmerpenguins::penguins,
  ggplot2::aes(x = species, y = body_mass_g)
) +
  ggplot2::geom_boxplot(na.rm = TRUE) +
  ggplot2::theme_minimal(base_size = 12)
```

A hand-written div carries no prefix to strip, so the filter leaves it alone and Quarto numbers it in the same sequence.
It is the form for content that is neither a table nor a figure.

::: {#anx-notes}
Regression diagnostics, model assumptions and any other supporting material go here as ordinary blocks.

Free-form annexe written as a fenced div.
:::

# Cross-references

- Section reference: @sec-figure
- Figure reference: @fig-penguins
- Equation reference: @eq-ols
- Annexe reference: @anx-notes

\ 

# Session {.unnumbered .unlisted}

```{r}
#| label: session-info
#| echo: true

sessioninfo::session_info(pkgs = "attached")
```
