---
title: "hebstr-doc — Example Document"
subtitle: "Multi-format Quarto extension"
author: "hebstr"
date: last-modified
format:
  hebstr-doc-html:
    code-tools:
      source: https://github.com/hebstr/quarto-hebstr-doc
---

# Typography and headings

This document exercises every capability declared in `_extension.yml` so a single `quarto render example.qmd` validates the extension. Only `hebstr-doc-html` is declared on `main`; Typst and DOCX outputs will be added once those formats are validated.

The body font is **Luciole** (sans-serif, designed for low-vision readers). Inline `code()` uses **Fira Code** with programming ligatures.

## Level 2 heading

Headings are numbered down to level 4 (`number-depth: 4`). The cross-reference title delimiter is a literal dot (e.g. *Figure 1.1*).

### Level 3 heading

Paragraph with *emphasis*, **strong emphasis**, a [link to the Quarto docs](https://quarto.org), and an internal reference to @sec-figure.

#### Level 4 heading

Deepest numbered level.

# Lists

Unordered:

- First item
- Second item with a *nested* block:
    - Nested bullet one
    - Nested bullet two
- Third item

Ordered:

1. Initialise the R session
2. Load the tidyverse
3. Render the document

# Figure {#sec-figure}

@fig-penguins is generated inline by **ggplot2** using the `palmerpenguins` dataset, driven by the shared knitr chunk options from `_extension.yml` (`echo: false`, `message: false`, `warning: false`).

```{r}
#| label: fig-penguins
#| fig-cap: "Bill length vs flipper length by species (palmerpenguins)."
#| fig-width: 6
#| fig-height: 4

library(ggplot2)

ggplot(
  palmerpenguins::penguins,
  aes(x = flipper_length_mm, y = bill_length_mm, colour = species)
) +
  geom_point(size = 3, alpha = 0.85) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
  labs(
    x = "Flipper length (mm)",
    y = "Bill length (mm)",
    colour = "Species"
  ) +
  theme_minimal(base_size = 12)
```

The HTML output wraps figures in lightbox; click the image to enlarge.

# Table

Code chunk below is a simple aggregation rendered as a Markdown table. `tbl-cap-location: top` puts the caption above the table; `tbl-title: "Table"` is language-neutral.

```{r}
#| echo: true
#| code-fold: false

library(dplyr)

palmerpenguins::penguins |>
  count(species) |>
  rename(Sp = species)
```

# Code block

The chunk below is shown with `echo: true` (overriding the default) to exercise syntax highlighting, code folding, and the copy-on-hover button (HTML), plus the `code.tmTheme` TextMate theme applied by Typst via `#set raw(theme: …)` in `template.typ`.

```{r}
#| echo: true
#| eval: false

library(dplyr)

palmerpenguins::penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species, island) |>
  summarise(
    n = n(),
    mean_mass_g = mean(body_mass_g),
    .groups = "drop"
  ) |>
  arrange(desc(mean_mass_g))
```

# Code window

The blocks below exercise the [`mcanouil/quarto-code-window`](https://github.com/mcanouil/quarto-code-window) filter declared in the YAML front matter.
The extension configures `style: default` (plain filename header, no traffic lights or window chrome — see `_extension.yml`) ; with `auto-filename: true`, blocks without an explicit `filename=` show the language name as the header label.

## Auto-generated filename (language name)

```python
def greet(name: str) -> str:
    return f"Hello, {name}!"
```

```r
penguins |>
  dplyr::count(species)
```

```bash
quarto add hebstr/quarto-hebstr-doc --no-prompt
quarto render example.qmd
```

# Injected script

The block below is **not** an inline chunk — it is the content of `scripts/demo_penguins.R`, injected via the `script` shortcode (in-tree `filters/add-code-files.lua`). The shortcode reads the file, emits the populated code block with the `cell-code` class so Quarto's `code-fold` wraps it in a foldable `<details>` element (collapsed by default) ; clicking the filename header expands it. The `<summary>` text is rewritten to the filename by `filters/add-code-files.js` (derived from [`shafayetShafee/add-code-files`](https://github.com/shafayetShafee/add-code-files), MIT).

This is the canonical pattern for hebstr-doc: inline chunks display as `code-window`, externalised scripts display as foldable `script` blocks.

{{< script scripts/demo_penguins.R >}}

# Equation

Inline math: $y = \beta_0 + \beta_1 x + \varepsilon$.

Display math with a cross-reference target (@eq-ols):

$$
\hat{\beta} = (X^\top X)^{-1} X^\top y
$$ {#eq-ols}

The HTML output uses `html-math-method: plain` (plain CSS rendering, no MathJax).

# Callouts

Callouts use `callout-appearance: simple`.

::: {.callout-note}
# A note callout.
:::

::: {.callout-note}
# Le note callout.

{{< lipsum 1 >}}

:::

::: {.callout-tip}
# A tip callout.

{{< lipsum 1 >}}

:::

::: {.callout-caution}
# A caution callout.

{{< lipsum 1 >}}

:::

::: {.callout-warning}
# A warning callout.

{{< lipsum 1 >}}

:::

::: {.callout-important}
# A important callout.

{{< lipsum 1 >}}

:::

# Block quote

> A citation.
>
> --- author

# Cross-references

- Section reference: @sec-figure
- Figure reference: @fig-penguins
- Equation reference: @eq-ols

\ 

***

# Session {.unnumbered .unlisted}

```{r}
#| echo: true
sessioninfo::session_info(pkgs = "attached")
```
