---
title: "hebstr-doc"
subtitle: "Multi-format Quarto extension"
author: "hebstr"
format:
  hebstr-doc-html:
    code-tools:
      source: https://github.com/hebstr/quarto-hebstr-doc
---

```{r}
#| label: setup
#| include: false

source("_extensions/hebstr-doc/fonts/register.R")
```

# Typography and headings

Body text is set in [**Luciole**](https://luciole-vision.com), a typeface designed for readers with low vision.
Inline `code()` is set in [**Fira Code**](https://github.com/tonsky/FiraCode).

## Level 2 heading

Headings are numbered down to level 4.

### Level 3 heading

Paragraph with *emphasis*, **strong emphasis**, a [link to the Quarto docs](https://quarto.org), and an internal reference to @sec-figure.

#### Level 4 heading

Deepest numbered level.

# Lists

Unordered:

- First item
- Second item with a *nested* block:
  - Nested bullet one
  - Nested bullet two
- Third item

Ordered:

1. Initialise the R session
2. Load the tidyverse
3. Render the document

# Figure {#sec-figure}

@fig-penguins follows the colour scheme: the chunk produces one plot per scheme with `renderings: [light, dark]`, on a transparent background.
Switch the scheme with the toggle to see the other version.

::: {#fig-penguins}
```{r}
#| label: penguins-fig
#| renderings: [light, dark]
#| fig-height: 5
#| dev.args: !expr list(bg = "transparent")

library(ggplot2)

theme_hebstr <- function(ink, grid) {
  theme_minimal(
    base_size = 12,
    base_family = "Luciole",
    ink = ink,
    paper = NA
  ) +
    theme(
      axis.text = element_text(colour = ink),
      panel.grid = element_line(colour = grid),
      axis.ticks = element_line(colour = grid)
    )
}

p <- ggplot(
  palmerpenguins::penguins,
  aes(x = flipper_length_mm, y = bill_length_mm, colour = species)
) +
  geom_point(size = 3, alpha = 0.85) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
  labs(
    x = "Flipper length (mm)",
    y = "Bill length (mm)",
    colour = "Species"
  )

p + theme_hebstr(ink = "#1a1a1a", grid = "#d9d9d9")
p + theme_hebstr(ink = "#e8e8e8", grid = "#4d4d4d")
```

Bill length vs flipper length by species (palmerpenguins).
:::

A figure without a cross-reference label keeps its caption but gets no number.

```{r}
#| label: penguins-mass-histogram
#| fig-cap: "Body mass distribution (palmerpenguins)."
#| fig-height: 3

ggplot(palmerpenguins::penguins, aes(x = body_mass_g)) +
  geom_histogram(bins = 30, na.rm = TRUE) +
  labs(x = "Body mass (g)", y = "Count") +
  theme_minimal(base_size = 12, base_family = "Luciole")
```

# Table

Table captions are placed above the table.

## Markdown table

```{r}
#| label: penguins-count
#| echo: true
#| code-fold: false

library(dplyr)

palmerpenguins::penguins |>
  count(species) |>
  rename(Sp = species)
```

## gt

@tbl-penguins-gt sets a light palette in R.
In dark mode, the theme replaces it with its own colours.

```{r}
#| label: tbl-penguins-gt
#| tbl-cap: "Bill length and body mass by species and island."
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species, island) |>
  summarise(
    n = n(),
    bill = round(mean(bill_length_mm, na.rm = TRUE), 1),
    mass = round(mean(body_mass_g)),
    .groups = "drop"
  ) |>
  gt::gt() |>
  gt::cols_label(
    species = "Species",
    island = "Island",
    n = "n",
    bill = "Bill length (mm)",
    mass = "Body mass (g)"
  ) |>
  gt::tab_spanner(label = "Mean", columns = c(bill, mass)) |>
  gt::cols_align(align = "left", columns = c(species, island)) |>
  gt::cols_align(align = "center", columns = c(n, bill, mass)) |>
  gt::tab_source_note("palmerpenguins, records with a body mass.") |>
  gt::tab_options(
    table.width = gt::pct(80),
    table.background.color = "#F0FAFF",
    row.striping.include_table_body = TRUE,
    row.striping.background_color = "white",
    column_labels.background.color = "white",
    column_labels.font.weight = "bold",
    column_labels.border.top.style = "none",
    column_labels.border.bottom.color = "#333333",
    source_notes.background.color = "white",
    table.font.color = "#333333",
    table.border.top.style = "none",
    table.border.bottom.style = "none",
    table_body.border.top.color = "#333333",
    table_body.border.bottom.color = "#333333",
    table_body.hlines.style = "none",
    data_row.padding = gt::px(8)
  )
```

## reactable

This widget uses the theme's CSS variables in `reactableTheme()`, so it follows the colour scheme.
A widget without a theme gets the same colours by default.

```{r}
#| label: penguins-reactable
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  filter(!is.na(body_mass_g)) |>
  select(species, island, bill_length_mm, body_mass_g) |>
  reactable::reactable(
    searchable = TRUE,
    filterable = TRUE,
    striped = TRUE,
    highlight = TRUE,
    defaultPageSize = 5,
    theme = reactable::reactableTheme(
      color = "var(--bs-body-color)",
      backgroundColor = "var(--primary-surface)",
      stripedColor = "var(--primary-back)",
      borderColor = "color-mix(in srgb, var(--neutral), transparent 70%)"
    )
  )
```

# Code block

Code is hidden by default; this chunk sets `echo: true` to show it, with syntax highlighting and a copy button.

```{r}
#| label: penguins-summary
#| echo: true
#| eval: false

library(dplyr)

palmerpenguins::penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species, island) |>
  summarise(
    n = n(),
    mean_mass_g = mean(body_mass_g),
    .groups = "drop"
  ) |>
  arrange(desc(mean_mass_g))
```

Printed output of a chunk whose code stays hidden gets a `default` header, since `collapse: true` merges it into a block with no language.
With `collapse: false`, as in this chunk, it shows as plain output on the same dark surface.

```{r}
#| label: penguins-mass-summary
#| collapse: false

summary(palmerpenguins::penguins$body_mass_g)
```

# Code window

Code blocks carry a header, provided by the embedded [`code-window`](https://github.com/mcanouil/quarto-code-window) extension.
Without a `filename` attribute, the header shows the language.

```python
def greet(name: str) -> str:
    return f"Hello, {name}!"
```

```r
penguins |>
  dplyr::count(species)
```

```bash
quarto add hebstr/quarto-hebstr-doc --no-prompt
quarto render example.qmd
```

# Injected script

The `script` shortcode includes `scripts/demo_penguins.R` as a folded code block; click the filename to expand it.

{{< script scripts/demo_penguins.R >}}

# Tabset {#sec-tabs}

A `.panel-tabset` div creates the tabs.

::: {.panel-tabset}
## Adelie

```{r}
#| label: penguins-adelie
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::filter(species == "Adelie") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```

## Chinstrap

```{r}
#| label: penguins-chinstrap
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::filter(species == "Chinstrap") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```

## Gentoo

```{r}
#| label: penguins-gentoo
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::filter(species == "Gentoo") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```
:::

# File tree

The `filetree` shortcode lists this repository at render time.
With `mode=dynamic`, folders open and close on click.
Descriptions come from the `filetree.yml` file.

{{< filetree mode=dynamic >}}

# Equation

Inline math: $y = \beta_0 + \beta_1 x + \varepsilon$.

Display math with a cross-reference target (@eq-ols):

$$
  \hat{\beta} = (X^\top X)^{-1} X^\top y
$$ {#eq-ols}

Display math with text (@eq-text):

$$
  \frac{\text{upper text}}{\text{lower text}} \text{ beside text}
$$ {#eq-text}

# Footnote

Footnotes are collected at the end of the document.[^footnote-demo]

[^footnote-demo]: The footnote body.
    Markdown works here: *emphasis*, `code()`, and [links](https://quarto.org).

# Margin note

A `.column-margin` div places its content in the margin column, next to the paragraph it follows.
Body text is justified; margin content stays left-aligned.

::: {.column-margin}
A margin note, written as a `.column-margin` div, with enough text to wrap over several lines in the narrow column.
:::

# Callouts

Callouts use `callout-appearance: simple`.

::: {.callout-note}
# A note callout.

{{< lipsum 1 >}}
:::

::: {.callout-tip}
# A tip callout.

{{< lipsum 1 >}}
:::

::: {.callout-caution}
# A caution callout.

{{< lipsum 1 >}}
:::

::: {.callout-warning}
# A warning callout.

{{< lipsum 1 >}}
:::

::: {.callout-important}
# An important callout.

{{< lipsum 1 >}}
:::

# Block quote

> A citation.
>
> --- author

# Annexes

Annexes are numbered separately and referenced as `@anx-…`.
A table chunk labelled `tbl-anx-…`, a figure chunk labelled `fig-anx-…` and a fenced div labelled `anx-…` all produce an annexe.

```{r}
#| label: tbl-anx-mass
#| tbl-cap: "Mean body mass by species and island."
#| echo: true
#| code-fold: false

palmerpenguins::penguins |>
  dplyr::group_by(species, island) |>
  dplyr::summarise(
    n = dplyr::n(),
    mean_mass_g = round(mean(body_mass_g, na.rm = TRUE)),
    .groups = "drop"
  )
```

```{r}
#| label: fig-anx-mass-spread
#| fig-cap: "Body mass spread by species."
#| fig-height: 3.5
#| echo: true
#| code-fold: false

ggplot2::ggplot(
  palmerpenguins::penguins,
  ggplot2::aes(x = species, y = body_mass_g)
) +
  ggplot2::geom_boxplot(na.rm = TRUE) +
  ggplot2::theme_minimal(base_size = 12)
```

A fenced div holds content that is neither a table nor a figure.

::: {#anx-notes}
Regression diagnostics, model assumptions and other supporting material.

Free-form annexe written as a fenced div.
:::

# Cross-references

- Section reference: @sec-figure
- Figure reference: @fig-penguins
- Equation reference: @eq-ols
- Annexe reference: @anx-notes

\ 

# Session {.unnumbered .unlisted}

```{r}
#| label: session-info
#| echo: true

sessioninfo::session_info(pkgs = "attached")
```
